Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-29 colt-21/cvx-flats.tex
Making sure Lcvx in appendix as appropriate

  \emph{$\Y=\reals$: }
  Any $L \in \Lcvx_d$ satisfies the assumptions of~\cite{ioffe1969minimization}, so we may interchange subdifferentiation and expectation.
  Specifically, letting $\V_{u,p} = \{V:\Y\to\reals^d \mid V \text{ measurable}, V(y) \in \partial L(u,y) \text{ $p$-a.s.}\}$, we have
  $\partial \E_p L(u,Y) = \{\int V(y)dp(y) \mid V\in\V_{u,p}\}$.
  As $\vec 0 \in \partial \E_p L(u,Y)$, in particular, there is some $V_{u,p} \in \V_{u,p}$ such that $\E_p V_{u,p} = 0$.
  \jessiet{Fleshing this out: $\exists E \subseteq \mathcal{B}(\Y)$ so that $p(E) = 0$ and $V(y) \in \partial L(u,y) \forall y \in E^c$.  Claim $q(E) = 0$, which would complete the claim that $V(y) \in \partial L(u,y) q$-a.s.  For contradiction, if $q(E) > 0$, then $\exists x \in E$ and open neighborhood $N_x \ni x$ with $q(N_x) > 0$.  Since $\supp(p) = \supp(q)$, we then have $p(N_x) >0$, contradicting $p(E) = 0$, and hence $V(y) \in \partial L(u,y)$ not $p$-a.s.  Therefore, we must have $q(E) = 0$, so $V(y) \in \partial L(u,y) q$-a.s.}
  For any $q\in\P$, as by assumption $q$ is supported on the same set as $p$, we have $V_{u,p}(y)\in\partial L(u,y)$ $q$-a.s., so that $V_{u,p}\in\V_{u,q}$\raft{I'm a bit too tired to see exactly how this follows, but it seems intuitively obvious. I'm okay to punt on it for the submission, but if someone has time to check, that would be great!}.
  Thus, $\E_q V_{u,p} = 0$ implies $0\in\partial\E_q L(u,Y)$ by the above.
  \raft{For posterity, the bug: we want to conclude $\E_q V_{u,p} = 0 \implies q \in \Gamma_u$.  But the ``a.s.\ in $p$'' means we may not have $0 \in \partial \E_q L(u,Y)$, since maybe $V_{u,p}(y)$ is not in $\partial L(u,y)$ a.s.\ in $q$...}

@@

	As above, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L \in \L_d$ is consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

        \jessie{Adding an attempt at sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) = d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there should be some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, we can verify $\Sc_{\gamma_r}(p) + \{p\}$ is one such $d$-flat.
    If $p \in \relint{\simplex}$, then the feasible subspace  (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and in particular, we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-d$.

@@

	When $p \not \in \relint{\simplex}$, we can loosely ``project down'' to the subsimplex on the support of $p$ and modify $L$ and $\ell$ accordingly.
	Now $p$ is in the relative interior of this subsimplex, so the above result gives $\dim(\Sc_{\gamma_r}(p)) \geq \|p\|_0 - d - 1$, where the feasible subspace dimension is relative to $\reals^{\supp(p)}$.
	Finally, we observe that the feasible subspace dimension in the projected space is the same as in the original space because of $p$'s location on a face of $\simplex$.
\end{proof}

@@

@@

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width=.45\textwidth]{tikz/fsd-bound.pdf} &
		\includegraphics[width=.45\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
          (Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) > 1$.}
	\label{fig:fsd-flats-abstain}
\end{figure}

@@

%The proof of the following Theorem is relegated to Appendix~\ref{app:pf-bayesrisklowerbound}
\begin{restatable}{theorem}{bayesrisklowerbound}\label{thm:bayes-risk-lower-bound}
  Let $\P$ be a set of Lebesgue densities supported on the same set for all $p \in \P$.
  Let $\Gamma:\P\to\reals^d$ satisfy Condition~\ref{cond:v-interior} for some $r\in\reals^d$.
  Let $L$ elicit $\Gamma$ such that $\lbar$ is non-constant on $\Gamma_r$.
  Then $\conscvx(\lbar) \geq \eliccvx(\lbar) \geq d+1$.
\end{restatable}

@@

Moreover, we have calibration implying indirect elicitation.
\begin{lemma}\label{lem:calib-implies-indir}
	If a surrogate and link $(L, \psi)$ are calibrated with respect to a loss $\ell:\R \times\Y \to \reals$, then $L$ indirectly elicits the property $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	Let $\Gamma$ be the unique property directly elicited by $L$, and fix $p \in \simplex$ with $u$ such that $p \in \Gamma_u$.
	We know such a $u$ exists since $\Gamma(p) \neq \emptyset$.
	As $p \in \Gamma_u$, then $\zeta(\exploss{L}{u}{p} - \risk{L}(p)) = \zeta(0) = 0$, we observe the bound $\ell(\psi(u); p) \leq \risk{\ell}(p)$.
	We also have $\ell(\psi(u); p) \geq \risk{\ell}(p)$ by definition of $\risk{\ell}$, so we must have $\ell(\psi(u);p) = \risk{\ell}(p) = \ell(\gamma(p); p)$, and therefore, $p \in \gamma_{\psi(u)}$.
	Thus, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, so $L$ indirectly elicits $\gamma$.
\end{proof}

@@

\hariresult*
\begin{proof}
  Let $L : \reals^d \times \Y \to \reals$ elicit $\Gamma$ be a calibrated surrogate for $\ell$, and consider $\Y^{\proj(q)} := \supp(q)$ for any $q \in \simplex$ and what happens when we restrict $L$ and $\ell$ to only the outcomes in $\Y^{\proj(q)}$.
  \raft{I'm not getting the $\proj(q)$ stuff.  It would help if you defined $\proj(q)$, but even so, I suspect there is a more direct approach.  E.g. you define all these $\proj(q)$ versions and then just take the primed versions to be equal.  I'd suggest starting from scratch and thinking through what exactly you need.}
  Take $L^{\proj(q)} := L|_{\Y^{\proj(q)}}$ and $\ell^{\proj(q)} := \ell|_{\Y^{\proj(q)}}$.
  Observe that for all $q \in \simplex$, we have $\gamma(q) = \gamma^{\proj(q)}(B(q))$.

@@

\begin{lemma}[{\citet[Lemma 8]{frongillo2020elicitation}}]
  \label{lem:elic-complex-bayes-concave}
  Suppose $L$ elicits $\Gamma:\P\to\R$ and has Bayes risk $\lbar$.
  Then for any $p,p'\in\P$ with $\Gamma(p)\neq\Gamma(p')$, we have $\lbar(\lambda p + (1-\lambda) p') > \lambda \lbar(p) + (1-\lambda) \lbar(p')$ for all $\lambda\in(0,1)$.
\end{lemma}

>>>>

  \emph{$\Y=\reals$: }
  Any $L \in \Lcvx_d$ satisfies the assumptions of~\cite{ioffe1969minimization}, so we may interchange subdifferentiation and expectation.
  Specifically, letting $\V_{u,p} = \{V:\Y\to\reals^d \mid V \text{ measurable}, V(y) \in \partial L(u,y) \text{ $p$-a.s.}\}$, we have
  $\partial \E_p L(u,Y) = \{\int V(y)dp(y) \mid V\in\V_{u,p}\}$.
  As $\vec 0 \in \partial \E_p L(u,Y)$, in particular, there is some $V_{u,p} \in \V_{u,p}$ such that $\E_p V_{u,p} = 0$.
  \jessiet{Fleshing this out: $\exists E \subseteq \mathcal{B}(\Y)$ so that $p(E) = 0$ and $V(y) \in \partial L(u,y) \forall y \in E^c$.  Claim $q(E) = 0$, which would complete the claim that $V(y) \in \partial L(u,y) q$-a.s.  For contradiction, if $q(E) > 0$, then $\exists x \in E$ and open neighborhood $N_x \ni x$ with $q(N_x) > 0$.  Since $\supp(p) = \supp(q)$, we then have $p(N_x) >0$, contradicting $p(E) = 0$, and hence $V(y) \in \partial L(u,y)$ not $p$-a.s.  Therefore, we must have $q(E) = 0$, so $V(y) \in \partial L(u,y) q$-a.s.}
  For any $q\in\P$, as by assumption $q$ is supported on the same set as $p$, we have $V_{u,p}(y)\in\partial L(u,y)$ $q$-a.s., so that $V_{u,p}\in\V_{u,q}$\raft{I'm a bit too tired to see exactly how this follows, but it seems intuitively obvious. I'm okay to punt on it for the submission, but if someone has time to check, that would be great! \jessie{See above margin comment.  I think it's too much detail for the conference submission, but feel free to change my mind.}}.
  Thus, $\E_q V_{u,p} = 0$ implies $0\in\partial\E_q L(u,Y)$ by the above.
  \raft{For posterity, the bug: we want to conclude $\E_q V_{u,p} = 0 \implies q \in \Gamma_u$.  But the ``a.s.\ in $p$'' means we may not have $0 \in \partial \E_q L(u,Y)$, since maybe $V_{u,p}(y)$ is not in $\partial L(u,y)$ a.s.\ in $q$...}

@@

	As above, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L \in \Lcvx_d$ is consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

        \jessie{Adding an attempt at sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) \geq d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there is some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, we can verify $\Sc_{\gamma_r}(p) + \{p\}$ is one such $d$-flat.
    If $p \in \relint{\simplex}$, then the feasible subspace  (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-\dim(\Sc_{\gamma_r(p)})$.

@@

	When $p \not \in \relint{\simplex}$, we can loosely ``project down'' to the subsimplex on the support of $p$ and modify $L$ and $\ell$ accordingly.
	Now $p$ is in the relative interior of this subsimplex, so the above result gives $\conscvx(\gamma) \geq \|p\|_0 - 1 - \dim(\Sc_{\gamma_r}(p))$, where the feasible subspace dimension is relative to $\reals^{\supp(p)}$.
	Finally, we observe that the feasible subspace dimension in the projected space is the same as in the original space because of $p$'s location on a face of $\simplex$.
\end{proof}

@@

It's worth noting that bounds provided by $d$-flats appear to work well at distributions complimentary to those providing previous bounds via feasible subspace dimension.
In some rough sense, this is because flats allow us a ``global'' view of the property we are eliciting, while the feasible subspace method only permits a ``local'' look at the property, so we find our method works better for distributions in $\relint\simplex$.
\jessie{Can someone check this paragraph?}

@@

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width=.45\textwidth]{tikz/fsd-bound.pdf} &
		\includegraphics[width=.45\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
          (Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) geq 2$.}
	\label{fig:fsd-flats-abstain}
\end{figure}

@@

%The proof of the following Theorem is relegated to Appendix~\ref{app:pf-bayesrisklowerbound}
\begin{restatable}{theorem}{bayesrisklowerbound}\label{thm:bayes-risk-lower-bound}
  Let $\P$ be a set of Lebesgue densities supported on the same set for all $p \in \P$.
  Let $\Gamma:\P\to\reals^d$ satisfy Condition~\ref{cond:v-interior} for some $r\in\reals^d$.
  Let $L \in \Lcvx$ elicit $\Gamma$ such that $\lbar$ is non-constant on $\Gamma_r$.
  Then $\conscvx(\lbar) \geq \eliccvx(\lbar) \geq d+1$.
\end{restatable}

@@

Moreover, we have calibration implying indirect elicitation.
\begin{lemma}\label{lem:calib-implies-indir}
	If a surrogate and link $(L, \psi)$ with $L \in \L$ are calibrated with respect to a loss $\ell:\R \times\Y \to \reals$, then $L$ indirectly elicits the property $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	Let $\Gamma$ be the unique property directly elicited by $L$, and fix $p \in \simplex$ with $u$ such that $p \in \Gamma_u$.
	We know such a $u$ exists since $\Gamma(p) \neq \emptyset$.
	As $p \in \Gamma_u$, then $\zeta(\exploss{L}{u}{p} - \risk{L}(p)) = \zeta(0) = 0$, we observe the bound $\ell(\psi(u); p) \leq \risk{\ell}(p)$.
	We also have $\ell(\psi(u); p) \geq \risk{\ell}(p)$ by definition of $\risk{\ell}$, so we must have $\ell(\psi(u);p) = \risk{\ell}(p) = \ell(\gamma(p); p)$, and therefore, $p \in \gamma_{\psi(u)}$.
	Thus, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, so $L$ indirectly elicits $\gamma$.
\end{proof}

@@

\hariresult*
\begin{proof}
  Let $L \in \Lcvx_d$ elicit $\Gamma$ be a calibrated surrogate for $\ell$, and consider $\Y^{\proj(q)} := \supp(q)$ for any $q \in \simplex$ and what happens when we restrict $L$ and $\ell$ to only the outcomes in $\Y^{\proj(q)}$.
  \raft{I'm not getting the $\proj(q)$ stuff.  It would help if you defined $\proj(q)$, but even so, I suspect there is a more direct approach.  E.g. you define all these $\proj(q)$ versions and then just take the primed versions to be equal.  I'd suggest starting from scratch and thinking through what exactly you need.}
  Take $L^{\proj(q)} := L|_{\Y^{\proj(q)}}$ and $\ell^{\proj(q)} := \ell|_{\Y^{\proj(q)}}$.
  Observe that for all $q \in \simplex$, we have $\gamma(q) = \gamma^{\proj(q)}(B(q))$.

@@

\begin{lemma}[{\citet[Lemma 8]{frongillo2020elicitation}}]
  \label{lem:elic-complex-bayes-concave}
  Suppose $L \in \L$ elicits $\Gamma:\P\to\R$ and has Bayes risk $\lbar$.
  Then for any $p,p'\in\P$ with $\Gamma(p)\neq\Gamma(p')$, we have $\lbar(\lambda p + (1-\lambda) p') > \lambda \lbar(p) + (1-\lambda) \lbar(p')$ for all $\lambda\in(0,1)$.
\end{lemma}
